---
title: "STAT 461 - ANOVA/Post Hoc Report on Fruit Fly Longevity"
author: "Mitchell Wentzel, Mason Conrad, Luke Geier, Marquis Gee"
date: "March 31, 2022"
output: pdf_document
geometry: left=1in,right=1in,top=1in,bottom=1in
urlcolor: blue
header-includes:
  - \usepackage{subfig}
---

## Set-Up R Environment

In the below code chunk, we will be running all of the necessary setup functions needed to compute the ANOVA/Post Hoc Analysis for the longevity of fruitfly lifespans.

```{r setup, include = FALSE}
# Clean R Environment.
rm(list = ls())

# Setting Document Options.
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

# Add additional packages by name to the following list.
packages <- c(
  "tidyverse", "knitr", "kableExtra",
  "psych", "car", "parameters"
  ) 
lapply(X = packages, FUN = library, character.only = TRUE)
library(readr)

# Loading Helper Files and Setting Global Options.
options(knitr.kable.NA = "")
options("contrasts" = c("contr.sum", "contr.poly"))
source("https://raw.github.com/neilhatfield/STAT461/master/rScripts/ANOVATools.R")
source("https://raw.github.com/neilhatfield/STAT461/master/rScripts/shadowgram.R")
```

## Load in the Data Set

In the following code chunk, we will be loading in the data set by using the read.csv() function, and specifying that the data set has a header (which indicates that column names exist in our csv file).

```{r data}
# Load in the fruitflies.csv data set.
fruitflies <- read.csv("fruitflies.csv", header = TRUE)

# Produce a head of the data set, which will show us the first few rows and the
# contents of our data frame.
head(fruitflies)
```

As we can see from the data head above, we have two columns in our fruitflies data frame. We have a column called "longevity" and a column called "condition." The "longevity" column is the value for the total lifespan of the fruitfly, in days. The "condition" column is a factor value that tells us the mating condition of the fruitfly. 

# Introduction and Background

The following report will detail the full data analysis for an ANOVA and Post Hoc Analysis of the fruitfly data set. The lifespan of a fruitfly can range from very little days up to hundreds of days. While there are many factors that could determine the lifespan of a fruitfly, the mating condition of a fruitfly has been claimed to be potentially significant in determining how long a fruitfly remains alive. 

Scientists have been able to mate fruitflies based on 5 different mating conditions and recorded the total days the fruitflies remained alive after they mated. The 5 mating conditions are "Null", "OnePreg", "OneVirgin", "EightPreg", and "EightVirgin". With 126 records of fruitfly mating condition and longevity of remaining lifespan, that leads us to our SRQ...

Does the mating condition of a fruitfly significantly affect the longevity of their remaining lifespan (in days)?

# Study Design and Methods

Here, write the narrative format of the study design-that is everything we've discussed in Unit 2. Incorporate the Hasse diagram, your hypotheses for any SRQs, any sample size considerations, as well as how you are going to handle the multiple comparison problem.

Include access link to the data set source.

## Hasse Diagram

<!-- 

{r hasse}
modelLabels <- c("1 Mate fruitflies 1", "5 (condition of mates) 4", "126 (male fruitflies) 121")
modelMatrix <- matrix(
  data = c(FALSE, FALSE, FALSE, TRUE, FALSE, FALSE, TRUE, TRUE, FALSE),
  nrow = 3,
  ncol = 3,
  byrow = FALSE
)
hasseDiagram::hasse(
 data = modelMatrix,
 labels = modelLabels
)


-->

Briefly assess the results of the code above.

If applicable, this is a subsection where you describe the sample/population that you will be working with along demographic lines.

# Exploration of the Data

Complete EDA here. Include visualizations and descriptive statistics and explaion the group's thoughts through looking at each result.

```{r}
ggplot(fruitflies, aes(x=longevity)) + geom_histogram() + xlab("Longevity") + ylab("Count")
```

Explanation here

```{r}
ggplot(fruitflies, aes(x=condition, y=longevity)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=3)
```

Explanation here

```{r}
shadowgram(
  dataVec = fruitflies$longevity,
  label = "Longevity",
  layers = 50,
  color = "navy",
  aStep = 4
)
```

Group discussion here.

# Results

This is the section of the report where you'll carry out an inferential methods.

```{r}
fruitflies$condition <- as.factor(fruitflies$condition)

flyModel <- aov(
  formula = longevity ~ condition,
  data = fruitflies,
  na.action = "na.omit"
)
```

## Assumptions

As this subsection's title implies, you'll discuss any assessment of assumptions in this section BEFORE you share any results for the related methods.

```{r}
car::qqPlot(
x = flyModel$residuals,
distribution = "norm",
envelope = 0.90,
id = FALSE,
pch = 20,
ylab = "Residuals"
)
```

The qq-plot 

```{r}
ggplot(
data = data.frame(
residuals = flyModel$residuals,
fitted = flyModel$fitted.values
),
mapping = aes(x = fitted, y = residuals)
) +
geom_point(size = 2) +
theme_bw() +
xlab("Fitted values (days)") +
ylab("Residuals (days)")

```

The strip chart

```{r}
ggplot(
data = data.frame(
residuals = flyModel$residuals,
index = 1:length(flyModel$residuals)
),
mapping = aes(x = index, y = residuals)
) +
geom_point(size = 1.5) +
geom_line() +
theme_bw() +
geom_hline(
yintercept = 0,
linetype = "dashed",
color = "red"
) +
xlab("Measurement order") +
ylab("Residuals")

```

The index plot

2-3 paragraphs, group discussion of the results of the assumptions.

## Omnibus  

MAKE A DECISION

Parametric Case

```{r parametricCase}
parameters::model_parameters(
  model = flyModel,
  omega_squared = "raw",
  eta_squared = "raw",
  epsilon_squared = "raw"
) %>%
  knitr::kable(
  digits = 4,
  col.names = c(
    "Source", "SS", "df", "MS", "F", "p-value",
    "Omega Sq.", "Eta Sq.", "Epsilon Sq."), 
  caption = "ANOVA Table for Fly Longevity Study",
  booktabs = TRUE,
  align = c("l", rep("c", 8))
  ) %>%
  kableExtra::kable_styling(
    font_size = 10,
    latex_options = c("HOLD_position")
  )

```

Our p-value of 0 is less than our unusualness threshold of 0.03, therefore we reject the null hypothesis and conclude that the mating condition of a fruitfly does have a statistically significant impact on the longevity of their remaining lifespan (in days).

Now we will calculate and display the point estimates for each condition.

```{r pointEstimates}
pointEst <- dummy.coef(flyModel)
pointEst <- unlist(pointEst)

data.frame("Estimate" = pointEst) %>%
  knitr::kable(
  digits = 2,
  caption = "Point Estimates from the Fly Longevity Study",
  format = "latex",
  booktabs = TRUE,
  align = "c"
  ) %>%
  kableExtra::kable_styling(
    font_size = 12,
    latex_options = c("HOLD_position")
  ) 

```

Discuss the above results, 1 paragraph.

## Post Hoc

Introductory sentence or two for post hoc.

MAKE A DECISION

Parametric Case

```{r tukeyHSD}
hsdFLY <- TukeyHSD(
  x = flyModel, # Your aov/lm object
  conf.level = 0.94 # 1 -- Your overall Type I Error level
)

## Kable Code for Tukey HSD
knitr::kable(
  x = hsdFLY$condition, # Notice the factor's name
  digits = 3,
  caption = "Post Hoc Tukey HSD Comparisons",
  col.names = c("Difference", "Lower Bound",
                "Upper Bound", "Adj. p-Value"),
  align = 'cccc',
  booktabs = TRUE,
) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("condensed", "boardered"),
    font_size = 12,
    latex_options = "HOLD_position"
  )

```

Discuss the above results, 1-2 paragraphs.

```{r postHocES1}
anova.PostHoc(flyModel) %>%
  knitr::kable(
    digits = 3,
    caption = "Post Hoc Comparison Effect Sizes",
    col.names = c("Pairwise Comparison","Cohen's d", "Hedge's g",
                  "Prob. Superiority"),
    align = 'lccc',
    booktabs = TRUE
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("condensed", "boardered"),
    font_size = 12,
    latex_options = "HOLD_position"
  )

```

# Limitations

As with any other statistical research analysis, our group did notice some short-comings and limitations to our study. We mainly indicated that there are many additional attributes that could have been tracked in order to receive more accurate and thorough results through our ANOVA/Post Hoc Analysis. A few additional attributes that could be seen as limitations to our research would be:

+ How many days old was the fruitfly before the mating took place?
+ Instead of testing only 1 and 8 for pregnant and virgin, why not add in 1 through 8?
+ How many total fruitflies were in the offspring of each instance of mating?

These are just a few of the limitations in the data that we noticed in our group. Adding these additional attributes could be beneficial to future studies regarding similar fruitfly data. 

\newpage

# Author Contributions

The authors of this report would like to acknowledge their individual contributions to the report.

+ Mitchell Wentzel contributed to the design of the study, collection of data, analysis of data, and writing of the report.
+ Mason Conrad contributed to the design of the study, collection of data, analysis of data, and writing of the report.
+ Luke Geier contributed to the design of the study, collection of data, analysis of data, and writing of the report.
+ Marquis Gee contributed to the design of the study, collection of data, analysis of data, and writing of the report.

\newpage

# Code Appendix

```{r codeAppendix, ref.label = knitr::all_labels(), echo = TRUE, eval = FALSE}

```
